---
title: "Stat/CS 187: Homework 4"
author: "yourname"
date: "Spring 2022"
output: 
  html_document:
      toc: true
      df_print:  paged 
---

  
When creating your Rmd script, be sure that all code will show up in the knitted document. Once you get your code working, do suppress messages and warnings -- only as needed (i.e., don't just remove them for every code chunk)
  

```{r setup}
knitr::opts_chunk$set(echo = TRUE,
                      fig.width=8, 
                      fig.height=5) 

# Load the packages we will be using
pacman::p_load(socviz, tidyverse, usmap, 
               maps, statebins, grid)


```  



## 1. Insurify DUI Graph

```{r Q1_Data}
dui <- tribble(
  ~ model,           ~ rate,    ~ specific,
  "RAM 2500",            45.3,       "Yes",
  "Chevyrolet S-10",     35.9,       "Yes",
  "BMW 4-Series",        31.7,       "Yes",
  "AUDI A4",             31.1,       "Yes",
  "Dodge Dakota",        30.5,       "Yes",
  "Subaru WRX",          30.0,       "Yes",
  "BMW 7-Series",        29.0,       "Yes",
  "Ford Ranger",         28.9,       "Yes",
  "Chevrolet Silverado", 28.3,       "Yes",
  "GMC Sierra",          28.0,       "Yes",
  "Nat'l Average",       17.9,        "No"
)
```



### Part 1a) Basic Graph

Create the basic graph shown in the word document. Save it as gg_dui and make sure it is shown below the codechunk.

```{r 1a_basicGraph}
gg_dui <- 
  ggplot(data = dui,
         mapping = aes(x = rate,
                       y = reorder(model, rate))) +
  
  geom_col(mapping = aes(fill = specific),
           show.legend = F) +
  
  labs(x = "Drivers per 1,000 with one or more DUI citations on their record",
       y = NULL,
       title = "iNSURiFY",
       subtitle = "Car Models with the Most DUIs",
       caption = "Source: Insurify analysis of over 2.7 million car insuarance applications") + 
  theme_grey()

gg_dui
```






### Part 1b) Theme and color choice

Add the additional changes by choosing the correct theme and specifying choices for the x & y-axis and color of the bars so they match the graph shown in 1b of the word document.

Save it as *gg_dui2*. Make sure that it appears below the code chunk in your knitted document

```{r 1b_intermediateGraph}
gg_dui2 <- 
  gg_dui +
  
  theme_test() +
  
  scale_x_continuous(breaks = seq(0,50,5),
                     expand = expansion(mult = 0,
                                        add = c(0, 5))) +
  
  scale_fill_discrete(type = c(Yes = "#660033",
                               No = "#6699CC")) 

gg_dui2
```




### Part 1c) Use theme to alter non-data ink

Use different arguments in theme() to change the non-data ink options to create the plot shown in the word document.

```{r 1c_finalGraph}
gg_dui2 +
  
  geom_vline(xintercept = 0) +
  
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.ticks = element_blank(),
        
        plot.title = element_text(size = 12, 
                                  color = "#FF6633", 
                                  face = "bold"),
        
        plot.subtitle = element_text(size = 18, 
                                     face = "bold",
                                     hjust = 0.5),
        
        
        axis.title.x = element_text(size = 10,
                                    face = "bold"),
        
        plot.caption = element_text(face = "italic"))
```









## 2. Mapping income and unemployment

Set Up Your Project and Load Libraries

We'll use income and unemployment data from the USDA for Question 2. 

Read in the data files USDA2019.csv and USDA2019state.csv

It would be a good idea to clear out your global environment before starting question 2!

```{r Q2_setup}
# Read in data files:
usda_c <- read.csv('USDA_2019_County.csv')
usda_s <- read.csv('USDA_2019_State.csv')


# A map theme provided by our book:
theme_map <- function(base_size=9, base_family="") {
    require(grid)
    theme_bw(base_size=base_size, base_family=base_family) +
        theme(axis.line=element_blank(),
              axis.text=element_blank(),
              axis.ticks=element_blank(),
              axis.title=element_blank(),
              panel.background=element_blank(),
              panel.border=element_blank(),
              panel.grid=element_blank(),
              panel.spacing=unit(0, "lines"),
              plot.background=element_blank(),
              legend.justification = c(0,0),
              legend.position = c(0,0),
              plot.title = element_text(hjust = 0.5)
              )
}

```



### 2a) Simple US Map

Using plot_usmap(), make the map shown in the word document. 

Try to match colors reasonably well, and remove the legend title.

```{r 2a_simplemap}
#  Use package usmap, and employment data 
plot_usmap(data = usda_s,
           regions = 'states',
           values = 'Median_Household_Income_2019',  # variable to graph
           color = 'white') +
  
  scale_fill_continuous(low = 'gold',
                        high = 'darkblue') +
  
  labs(title = "2019 Median Household Income") +
  
  theme(legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5, 
                                  size = 14))
```




### 2b) For more serious maps: Merge state lines info with USDA data.

Now, prepare to use the more 'serious' method of making a US map.  

First, save the data frame that contains the boundary lines for each state as we did in chapter 7.

Then merge the state USDA income data with map data you just created.

```{r 2b_mergedata}
us_states <- map_data("state")

us_states_usda <- 
  usda_s %>% 
  mutate(region = tolower(state))
```




### 2c) Map a Numeric variable by State: Median Income in 2019

Create the map shown in the word doc, matching the theme, title, legend label (none), font size, colors (colors don't have to be perfect). 

```{r 2c_mapnumeric}
ggplot(data = us_states_usda,
       mapping = aes(x = long, 
                     y = lat, 
                     group = group, 
                     fill = Median_Household_Income_2019)) +
  
  geom_polygon(color = 'gray90', 
               size = 0.2) +
  
  coord_map(projection = "albers", 
            lat0 = 39, 
            lat1 = 45)  +
  
  theme_map() + 
  
  labs(title = '2019 Median Household Income',
       fill = NULL) +
  
  scale_fill_gradient(low = 'slategray1', 
                      high = 'green4') +
  
  theme(title = element_text(hjust = 0.5, 
                             size = 14,
                             face = "bold"))
```

### 2d) Now map data at a county level. First, the data, merged with county map info

Merge the USDA county data with the county map data, as we did in chapter 7.  You will need to make a variable common to both files, in order to merge. To do this, use the data frame county_map, in the socviz package.  Don’t change county_map, but use it to create a new data frame with a variable called FIPS.  Since FIPS is numeric in the USDA data, you will need to make FIPS numeric in the new map data, too.

```{r 2d_countydata}
# county_map is a data frame in socviz -- do help(county_map) to see more
county_map2 <- county_map %>%  mutate(FIPS = as.numeric(id))

# merge the two together
county_usda <- county_map2 %>% left_join(usda_c, by = "FIPS")

# Print the first 6 rows of the resulting data set
head(county_usda)

```

### 2e) Map Median Household Income by county:

Map income by county, matching the map below, in terms of colors (pretty close is fine), title, legend label, font size, theme, etc.

```{r 2e_countymap}
ggplot(data = county_usda,
       mapping = aes(x = long, 
                     y = lat)) +
  geom_polygon(mapping = aes(group = group,  
                             fill = Median_Household_Income_2019)) +
  
  coord_equal() +
  
  scale_fill_gradient(low = 'lightcyan', 
                      high = 'darkblue') +
  
  theme_map() +
  
  labs(title = '2019 Median Income',  
       fill = 'Income')  +
  
  theme(title = element_text(size = 14,
                             face = "bold"))
```

