---
title: 'Stat/CS 187: Homework 4'
author: "yourname"
date: "Spring 2022"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    df_print: paged
---
  


```{r setup, message = FALSE}

## Set the default size of figures
knitr::opts_chunk$set(echo = TRUE) 

knitr::opts_chunk$set(echo = TRUE,
                      fig.width=8, 
                      fig.height=5) 

# Load the packages we will be using
pacman::p_load(hexmode,ggthemes) 

library(dplyr)
library(ggplot2)
library(ggthemes)
library(readr)

```  



## 1. Insurify DUI Graph

```{r Q1_Data}
dui <- tribble(
  ~ model,           ~ rate,    ~ specific,
  "RAM 2500",            45.3,       "Yes",
  "Chevyrolet S-10",     35.9,       "Yes",
  "BMW 4-Series",        31.7,       "Yes",
  "AUDI A4",             31.1,       "Yes",
  "Dodge Dakota",        30.5,       "Yes",
  "Subaru WRX",          30.0,       "Yes",
  "BMW 7-Series",        29.0,       "Yes",
  "Ford Ranger",         28.9,       "Yes",
  "Chevrolet Silverado", 28.3,       "Yes",
  "GMC Sierra",          28.0,       "Yes",
  "Nat'l Average",       17.9,        "No"
)
```



### Part 1a) 

```{r 1a_basicGraph}
# plot gg_dui as basic plot
gg_dui<-
  ggplot(data=dui,
         mapping=aes(x=rate,
                     y=reorder(model,rate),  # reorder 
                     fill=specific))+   # different color for groups
  geom_col(show.legend = FALSE)+     # hide legend for specific
# add title, subtitle,xlab and caption, also remove ylab  
  labs(title= "ISURIFY",
       subtitle = "Car Models with the Most DUIs",
       x= "Drivers per 1,000 with one or more DUIs citations on their record",
       y="",
       caption = "Source: Insurify analysis of over 2.7 million car insuarance applications")


  
  
# Making sure your knitted document shows the resulting plot
gg_dui
```






### Part 1b) 

```{r 1b_intermediateGraph}
# choose color
color_manual<-c("#6699CC","#660033")

# plot gg_dui2 based on gg_dui
gg_dui2 <- 
  gg_dui + theme_test()+    # change theme
# set scales and change breaks for x-axis
  scale_x_continuous(breaks=c(0,5,10,15,20,25,30,35,40,45,50), 
                     expand = expansion(mult = 0,
                                        add = c(0, 5)))+
# apply color designed into groups for dui$specific
  scale_fill_manual(values=color_manual)


  
# show gg_dui2
gg_dui2
```




### Part 1c) 

```{r 1c_finalGraph}
# change theme for non-data ink options
gg_dui2 + theme_tufte(ticks=FALSE)+
# add vertical line at x=0
  theme(axis.line.y = element_line(size=1,colour="black",linetype=1),
# change color & size of subtitle and size & position of subtitle
        plot.title= element_text(size=12,colour="#FF6633"),
        plot.subtitle = element_text(size=20,hjust=0.5))
        
  


```









## 2. Mapping income and unemployment

```{r Q2_setup}
# Read in data files:
US_2019_C <- read_csv("USDA_2019_County.csv")
US_2019_S <- read_csv("USDA_2019_State.csv")


## Set the default size of figures
knitr::opts_chunk$set(fig.width=8, fig.height=5)

# Changing the default theme
theme_set(theme_bw())

# Load the libraries we will be using
pacman::p_load(gapminder,socviz, tidyverse, grid,usmap, maps, statebins, viridis, leaflet)


# A map theme provided by our book:
theme_map <- function(base_size=9, base_family="") {
    require(grid)
    theme_bw(base_size=base_size, base_family=base_family) %+replace%
        theme(axis.line=element_blank(),
              axis.text=element_blank(),
              axis.ticks=element_blank(),
              axis.title=element_blank(),
              panel.background=element_blank(),
              panel.border=element_blank(),
              panel.grid=element_blank(),
              panel.spacing=unit(0, "lines"),
              plot.background=element_blank(),
              legend.justification = c(0,0),
              legend.position = c(0,0)
              )
}


```



### 2a) Simple US Map

```{r 2a_simplemap}

#  Use package usmap, and employment data 
plot_usmap(regions = c('state'),
           data = data.frame(US_2019_S),
           values= "Median_Household_Income_2019") +
# choose color reasonably close
  scale_fill_continuous(low = 'gold1', high = 'orchid4')+
# add title at center and remove legend title
  labs(title="2019 Median Household Income") + 
  theme(plot.title = element_text(size=15,hjust=0.5),
        legend.title = element_blank())




```




### 2b) For more serious maps: Merge state lines info with USDA data.

```{r 2b_mergedata}
# create state_lines using the map_data function like we did in class
state_lines <- map_data("state")
head(state_lines)
# clean data and change region to state for merge
state_lines<- state_lines%>%
  mutate(state=region)%>%
  select(-region,-subregion)

# change state to all lowercase
US_2019_S$state<-tolower(US_2019_S$state)

# merge data
US <- US_2019_S %>% 
  left_join(y = state_lines,
            by = 'state')

```




### 2c) Map a Numeric variable by State: Median Income in 2019

```{r 2c_mapnumeric}


ggplot(data = US,
         mapping = aes(x = long, 
                       y = lat, 
                       group = group))  +
  
  geom_polygon(aes(fill = US$Median_Household_Income_2019),color="white")  + 
  
  theme_map() + 
  coord_map(projection = "albers", 
            lat0 = 39, lat1 = 45)+
  
# choose color reasonably close  
  scale_fill_gradient(low = 'lightblue2', high = 'green4' )+
  
# add title at center and remove legend title
  labs(title="2019 Median Household Income") + 
  
  theme(plot.title = element_text(size=15,hjust=0.5),
        legend.title = element_blank())

```

### 2d) Now map data at a county level. First, the data, merged with county map info

```{r 2d_countydata}
# county_map is a data frame in socviz -- do help(county_map) to see more
?county_map

#create a new data frame using county_map data
new_cmap<-county_map%>% 
  select_all()%>%
# create a variable called FIPS
  mutate(FIPS = id)%>%
# drop id
  select(-id)
# change FIPS to numeric values
new_cmap$FIPS <- as.numeric(new_cmap$FIPS)


# merge the two together
US2 <- US_2019_C %>% 
  left_join(y = new_cmap,
            by = 'FIPS')

# Print the first 6 rows of the resulting data set
head(US2)



```

### 2e) Map Median Household Income by county:

**Map income by county, matching the map below, in terms of colors (pretty close is fine), title, legend label, font size, theme, etc.**

```{r 2e_countymap}
US2$Income<-US2$Median_Household_Income_2019
ggplot(data = US2,
         mapping = aes(x = long, y = lat,
                       fill=Income,
                       group = group
                       ))+
  
  geom_polygon(size=0.05)+
  theme_map()+  
# choose color reasonably close  
#  scale_fill_continuous(low="white",high="blue")+
  scale_fill_gradient(low="lightcyan",high="darkblue",na.value="gray")+
# add title at center and remove legend title
  labs(title="2019 Median Income") + 
  theme(plot.title = element_text(size=15,hjust=0.5),
        legend.title=element_text(size=15))


```

